On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-04-02 manuscripts/PNAS/chazthm_pnas.tex

\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix integers $n, \ell$, and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k-1){m \choose k}+1$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
\begin{align}\label{Cstable}
\|\mathbf{A}_j - (\mathbf{B}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon,\ \ \   \text{for  $j \in [m]$},
\end{align}
for some permutation matrix $\mathbf{P}$ and invertible diagonal $\mathbf{D}$.

Furthermore, if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$, then $\mathbf{B}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$, and:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,\ \ \   \text{for $i \in [N]$}.
\end{align}
\end{theorem}

@@

\begin{lemma}\label{NonEmptyLemma} 
Fix integers $\ell, m$ and $\hat m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular and satisfies the singleton intersection property. If there exists a map $\pi: E \to 2^{[\hat m]}$ such that $\sum_{S \in E} |S| = \sum_{S \in E} |\pi(S)|$ and:
\begin{align}\label{cond}
|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for all } F \subseteq E,
\end{align}
%
then $\hat m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $[J]$ to $[\hat m]$ for some $J \subseteq [m]$ of size $\hat m - \ell(\hat m - m)$. In particular, if $\hat m = m$ then the map $\pi$ is induced by a permutation.
%$\pi(E)$ is $\ell$-regular and satisfies the singleton intersection property. In particular, $|\cap_{S \in F} S| = 1$ if and only if $|\cap_{S \in F} \pi(S)| = 1$. 
\end{lemma}
\begin{proof}
[ ** TODO ** rewrite for $\hat m > m$. ]
We first show that $\hat m \geq m$. Fix $T = \{(j, S): j \in \pi(S), S \in E\}$. Then $|T| = \sum_{S \in E} |\pi(S)| = \sum_{S \in E} |S| = \sum_{i \in [m]} \deg(i) = \ell m$ by regularity of $E$. If $\hat m < m$ then by pigeonholing the $m \ell \geq \ell (\hat m + 1) \geq [(\ell + 1) - 1] \hat m + 1$ elements of $T$ with respect to the set $[\hat m]$ of possible first indices, we see that there must exist at least $\ell + 1$ elements of $T$ having the same first index. But by \eqref{cond} there can be no more than $\ell$ elements of $T$ with a given first index, since $E$ is $\ell$-regular. Hence $\hat m \geq m$. Again, pigeonholing the elements of $T$ with respect to the set $[\hat m]$ of possible first indices, we see that for each $j$ there must be exactly $\ell$ elements of $T$ with $j$ as a first index; hence, $\pi(E)$ is $\ell$-regular.

>>>>

\begin{theorem}\label{DeterministicUniquenessTheorem}
%Fix integers $n, \ell$, and $k < m$. 
Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} for some $k < m$ and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k-1){m \choose k}+1$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  Then there exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{\hat m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ must have $\hat m \geq m$ and, provided $p = \hat m - \ell(\hat m - m)$ is positive, necessarily satisfy:
\begin{align}\label{Cstable}
\|(\mathbf{A}_J- \mathbf{B}_{\hat J} \mathbf{PD} )_j\|_2 \leq C_1 \varepsilon \ \  \text{for all $j \in [J]$},
\end{align}
for some $J \in {[m] \choose p}$, $\hat J \in {[\hat m] \choose p}$, permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$.

[ ** TODO ** write $\hat m$ version ] Furthermore, if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$, then $\mathbf{B}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$, and:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,\ \ \   \text{for $i \in [N]$}.
\end{align}
\end{theorem}

%\begin{theorem}\label{DeterministicUniquenessTheorem}
%Fix integers $n, \ell$, and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k-1){m \choose k}+1$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

%Every matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
%\begin{align}\label{Cstable}
%\|\mathbf{A}_j - (\mathbf{B}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon,\ \ \   \text{for all $j \in [m]$},
%\end{align}
%for some permutation matrix $\mathbf{P}$ and invertible diagonal $\mathbf{D}$.

%Furthermore, if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$, then $\mathbf{B}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$, and:
%\begin{align}\label{b-PDa}
%\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,\ \ \   \text{for all $i \in [N]$}.
%\end{align}
%\end{theorem}

@@

\begin{lemma}\label{NonEmptyLemma} 
Fix integers $\ell, m$ and $\hat m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular and satisfies the singleton intersection property. If there exists a map $\pi: E \to 2^{[\hat m]}$ such that $\sum_{S \in E} |S| = \sum_{S \in E} |\pi(S)|$ and:
\begin{align}\label{cond}
|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for all } F \subseteq E,
\end{align}
%
then $\hat m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $[J]$ to $[\hat m]$ for some $J \subseteq [m]$ of size $\hat m - \ell(\hat m - m)$. In particular, if $\hat m = m$ then the map $\pi$ is induced by a permutation.
%$\pi(E)$ is $\ell$-regular and satisfies the singleton intersection property. In particular, $|\cap_{S \in F} S| = 1$ if and only if $|\cap_{S \in F} \pi(S)| = 1$. 
\end{lemma}
\begin{proof}
We first show that $\hat m \geq m$. Fix $T = \{(j, S): j \in \pi(S), S \in E\}$. Then $|T| = \sum_{S \in E} |\pi(S)| = \sum_{S \in E} |S| = \sum_{i \in [m]} \deg(i) = \ell m$ by regularity of $E$. If $\hat m < m$ then by pigeonholing the $m \ell \geq \ell (\hat m + 1) \geq [(\ell + 1) - 1] \hat m + 1$ elements of $T$ with respect to the set $[\hat m]$ of possible first indices, we see that there must exist at least $\ell + 1$ elements of $T$ having the same first index. But by \eqref{cond} there can be no more than $\ell$ elements of $T$ with a given first index, since $E$ is $\ell$-regular. Hence $\hat m \geq m$.

[ ** TODO ** rewrite for $\hat m > m$. ]
Again, pigeonholing the elements of $T$ with respect to the set $[\hat m]$ of possible first indices, we see that for each $j$ there must be exactly $\ell$ elements of $T$ with $j$ as a first index; hence, $\pi(E)$ is $\ell$-regular.
